A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-03 paper.tex
Shorter

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results.
The availability of resources only limited by your budget, and in many
cases for free, as well as the inherent reproducibility of cloud
deployments, has been a boon for scientific computing. However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer the best or most efficient architectural design
for scientific workflows and implementation of algorithms. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from a monolithic n-tier and mostly sequential
to an asynchronously parallel architecture will also imply important
reformulation in the algorithms to take full advantage of this new
mode.

  Cloud native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow that is managed by the
  cloud provider, allowing the user to focus on code and connections
  among different services. Describing them as {\em native} points
  to their new architecture, which departs from a monolithic or even
  distributed paradigm to become a loosely connected collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing and
  not keeping state from one invocation to the next.

  Reactive systems not only allow massive scaling and independent deployment they
  are also more economical than other monolithic options.  Platform as a service
  (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  and also for time they remain active.

  At any rate, while one of the main selling points of Functions as a Service is
  their ultra-fast activation time, from our point of view their most interesting
  feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps that work on a data
  stream.

    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    is the backbone of the application. The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be followed closely by releases by Azure
    (Microsoft) and Google and an open source implementation, OpenWhisk,
    released by IBM \cite{Baldini2016287}.

  These architectures are {\em event-based} in the sense that program
  flow is governed by events, which in this case are going to be simply
  messages received by the MessageHub service. These {\em events} trigger actions,
  and actions create new message/events which start the loop all over
  again. This is not the only mode of action in general serverless
  architectures; however, since they define a temporal axis and make new
  actions depend on the result of other actions, is the one that we are
  going to choose in this paper to implement a stateless version of an
  evolutionary algorithm.

>>>>

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.

  Cloud native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.

  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.

    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

  These architectures are {\em event-based} in the sense that program
  flow is governed by events.
  %, which in this case are going to be simply
  %messages received by the MessageHub service.
  %These {\em events} trigger actions,
  %and actions create new message/events which start the loop all over
  %again. This is not the only mode of action in general serverless
  %architectures; however, since they define a temporal axis and make new
  %actions depend on the result of other actions, is the one that we are
  %going to choose in this paper to implement a stateless version of an
  %evolutionary algorithm.
